Individual Variation in Animal Movement and Effects on Seed Dispersal Distance

### 2021-01-15 paper/paper.Rmd
updated figures and tables

```
\newpage
# Introduction
A main goal of ecologists and evolutionary biologists is to document and predict the structure and dynamics of ecological systems. Although most of our current understanding of ecological systems is based on the average patterns observed in nature, understanding variability and its consequences in ecological processes has recently raised the interests of researchers [@benedetti-cecchi_importance_2003; @inouye_importance_2005]. Variation is now considered a property of ecological systems that might contain as much ecological information as do averages and regularities in nature [@benedetti-cecchi_importance_2003; @violle_return_2012].
Individual variation, i.e. the possibility for individuals within a population to present different ecological strategies or traits [@violle_return_2012], has been widely discussed in ecology and evolutionary biology. In evolutionary biology the acknowledgement that individuals with different traits might vary in their fitness and survival under natural selection pressures dates back to Darwin’s seminal ideas [@darwin_origin_1859]. Today, it is widely recognized that individual variation plays a key role in the evolution of species and clades, and the relevance of individual variation in the performance of populations and species facing current threats (e.g. climate change, fragmentation and habitat loss) is a growing field of research [@forsman_inter-individual_2016]. In ecology, individual variation is often mentioned as a mechanism for populations to reduce competition pressure [@bolnick_ecological_2010; @araujo_ecological_2011]. Based on the niche complementarity principle, phenotypically different individuals will compete less strongly than more similar individuals [@bolnick_why_2011], thus individual variation would act as a release from intra- and interspecific competition at a local scale. Despite the well recognized importance of individual variation in ecological and evolutionary processes, ecological theory has emphasized interspecific variation [@violle_return_2012]. Not surprisingly, the role of individual variability on the outcome of species interactions has also been poorly studied; most models of species interactions assume that all conspecific individuals are equivalent and interchangeable [@bolnick_why_2011]. Individuals, however, differ in several aspects that will affect their interactions with other species. This variability makes individuals non-interchangeable from the perspective of interacting patterns [@bolnick_ecology_2003; @gonzalez-varo_labile_2016]. The consequences of individual variability in interspecies interactions can even have cascading effects, going beyond from the species-species level to the community and ecosystem level [@post_intraspecific_2008; @wolf_animal_2012].

@@

ggplot() +
  geom_point(data = seed_loc, aes(x = xloc, y = yloc), color = "red", size = 3) +
  geom_point(aes(x = 0, y = 0), color = "black", size = 5)+
  geom_vline(aes(xintercept=0), linetype = "dotted") + geom_hline(aes(yintercept = 0), linetype = "dotted") +
  #geom_segment(aes(x = seed_loc$xloc, xend = 0, y = seed_loc$yloc, yend = 0), linetype = "dashed", color = "red") +
  geom_point(data = mean_seed_loc, aes(x = x, y = y), color = "blue", size = 4) +
  geom_segment(aes(x = seed_loc$xloc, xend = mean_seed_loc$x, y = seed_loc$yloc, yend = mean_seed_loc$y), linetype = "dashed", color = "blue") +
  labs(x = "x", y = "y") +
  coord_fixed() +
  theme_bw() -> fig_agg

@@

We calculated seed dispersal distance as the euclidean distance of each seed to the parent plant. Given that our models only considered one parent plant located at the origin per simmulation, we calculated seed dispersal distance (DD) as the distance from each seed's location to the origin as follows:
$$DD_i = \sqrt{(0-x_i)^2 + (0-y_i)^2} $$
where $DD_i$ is the dispersal distance for seed $i$ in the simulation run and $x_i$ and $y_i$ are its xy coordinates in the landscape. We also used an aggregation metric to determine how evenly seeds would be dispersed across the landscape, and calculated seed dispersion (SD) as the average distance of each seed to the mean seed location in each simulation run (following methods in [@jones_closing_2017]):
$$SD_j = \frac{\sum \sqrt{(x_{mj} - x_{ij})^2 + (y_{mj} - y_{ij})^2}}{n}$$
where $n$ is the number of seeds for each simulation run (5), and $x_{mj}$ and $y_{mj}$ is the mean seed location in the $j$ simulation run. We obtained measures of seed dispersion for each simulation run, thus we calculated the average seed dispersion for each model as $\overline{SD} = \frac{\sum{SD_j}}{N}$, where $N$ is the total number of simulation runs for the model being considered.The information on seed dispersal distance was used for the toucan-generated seed dispersal kernels for each one of our models (null, individual, and family group), and these seed dispersal distances were also classified as long distance dispersal events if they exceeded 500m [@holbrook_using_2007]. The proportion of long-distance dispersal events, maximum dispersal distance, and average seed dispersion were used to compare seed dispersal between our three models of individual variation (Table1).

@@

## Seed dispersal distances and long distance dispersal

## Seed dispersion aggregation
The mean dispersal in the boxplots is not the same as the dispersal distances before. This mean dispersal is the distance from the origin to the mean seed location in each simulation run. The mean seed location is used to calculate dispersion, as the average distance of each seed to the mean seed location, as shown in Figure 1.

## Weibull seed dispersal kernel

>>>>

```{r}
load("paper/dispersal_kernels.RData")
```

\newpage
# Introduction
A main goal of ecologists and evolutionary biologists is to document and predict the structure and dynamics of ecological systems. Although most of our current understanding of ecological systems is based on the average patterns observed in nature, understanding variability and its consequences in ecological processes has recently raised the interests of researchers [@benedetti-cecchi_importance_2003; @inouye_importance_2005]. Variation is now considered a property of ecological systems that might contain as much ecological information as do averages and regularities in nature [@benedetti-cecchi_importance_2003; @violle_return_2012].
Individual variation, i.e. the possibility for individuals within a population to present different ecological strategies or traits [@violle_return_2012], has been widely discussed in ecology and evolutionary biology. In evolutionary biology the acknowledgement that individuals with different traits might vary in their fitness and survival under natural selection pressures dates back to Darwin’s seminal ideas [@darwin_origin_1859]. Today, it is widely recognized that individual variation plays a key role in the evolution of species and clades, and the relevance of individual variation in the performance of populations and species facing current threats (e.g. climate change, fragmentation and habitat loss) is a growing field of research [@forsman_inter-individual_2016]. In ecology, individual variation is often mentioned as a mechanism for populations to reduce competition pressure [@bolnick_ecological_2010; @araujo_ecological_2011]. Based on the niche complementarity principle, phenotypically different individuals will compete less strongly than more similar individuals [@bolnick_why_2011], thus individual variation would act as a release from intra- and interspecific competition at a local scale. Despite the well recognized importance of individual variation in ecological and evolutionary processes, ecological theory has emphasized interspecific variation [@violle_return_2012]. Not surprisingly, the role of individual variability on the outcome of species interactions has also been poorly studied; most models of species interactions assume that all conspecific individuals are equivalent and interchangeable [@bolnick_why_2011]. Individuals, however, differ in several aspects that will affect their interactions with other species. This variability makes individuals non-interchangeable from the perspective of interacting patterns [@bolnick_ecology_2003; @gonzalez-varo_labile_2016]. The consequences of individual variability in interspecies interactions can even have cascading effects, going beyond from the species-species level to the community and ecosystem level [@post_intraspecific_2008; @wolf_animal_2012].

@@

ggplot() +
  geom_point(data = seed_loc, aes(x = xloc, y = yloc), color = "red", size = 3) +
  geom_point(aes(x = 0, y = 0), color = "black", size = 5)+
  geom_vline(aes(xintercept=0), linetype = "dotted") + geom_hline(aes(yintercept = 0), linetype = "dotted") +
  #geom_segment(aes(x = seed_loc$xloc, xend = 0, y = seed_loc$yloc, yend = 0), linetype = "dashed", color = "red") +
  geom_point(data = mean_seed_loc, aes(x = x, y = y), color = "blue", size = 4) +
  geom_segment(aes(x = seed_loc$xloc, xend = mean_seed_loc$x, y = seed_loc$yloc, yend = mean_seed_loc$y),
               linetype = "dashed", color = "blue") +
  geom_segment(aes(x = 0, xend = mean_seed_loc$x, y = 0, yend = mean_seed_loc$y),
               linetype = "solid", color = "black") +
  labs(x = "x", y = "y") +
  coord_fixed() +
  theme_bw() -> fig_agg

@@

We calculated seed dispersal distance as the euclidean distance of each seed to the parent plant. Given that our models only considered one parent plant located at the origin per simulation, we calculated seed dispersal distance (DD) as the distance from each seed's location to the origin as follows:
$$DD_i = \sqrt{(0-x_i)^2 + (0-y_i)^2} $$
where $DD_i$ is the dispersal distance for seed $i$ in the simulation run and $x_i$ and $y_i$ are its xy coordinates in the landscape. We also used an aggregation metric to determine how evenly seeds would be dispersed across the landscape, and calculated seed dispersion (SD) as the average distance of each seed to the mean seed location in each simulation run (following methods in [@jones_closing_2017]):
$$SD_j = \frac{\sum \sqrt{(x_{mj} - x_{ij})^2 + (y_{mj} - y_{ij})^2}}{n}$$
where $n$ is the number of seeds for each simulation run (5), and $x_{mj}$ and $y_{mj}$ is the mean seed location in the $j$ simulation run. We obtained measures of seed dispersion for each simulation run, thus we calculated the average seed dispersion for each model as $\overline{SD} = \frac{\sum{SD_j}}{N}$, where $N$ is the total number of simulation runs for the model being considered.The information on seed dispersal distance was used for the toucan-generated seed dispersal kernels for each one of our models (null, individual, and family group), and these seed dispersal distances were also classified as long distance dispersal events if they exceeded 500m [@holbrook_using_2007]. The proportion of long-distance dispersal events, maximum dispersal distance, and average seed dispersion were used to compare seed dispersal between our three models of individual variation (Table1).

@@

## Seed dispersal distances and aggregation

```{r fig.width=6}
bind_rows(null_dispersal %>% mutate(model = "Null"), 
          indiv_dispersal %>% mutate(model = "Individual"),
          fam_dispersal %>% mutate(model = "Family")) %>% 
  mutate(model = factor(model, levels = c("Null", "Individual", "Family"))) -> dispersal_data

dispersal_data %>% 
  ggplot(., aes(x = dispersal, group = model)) +
  facet_wrap(~model, ncol = 3) +
  geom_histogram(aes(y = ..density..), binwidth = 50) +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white")) +
  labs(y = "Density", x = "Seed dispersal (m)") -> histograms

null_dispersion %>% 
  mutate(model = "Null") %>% 
  bind_rows(indiv_dispersion %>% 
              mutate(model = "Individual")) %>% 
  bind_rows(fam_dispersion %>% 
              mutate(model = "Family")) %>% 
  mutate(model = factor(model, levels = c("Null", "Individual", "Family"))) -> all_disp_data

all_disp_data %>% 
  ggplot(., aes(y = mean_xy_dispersal, x = model)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Mean dispersal (m)") +
  theme(axis.title.x = element_blank()) -> a

all_disp_data %>% 
  ggplot(., aes(y = seed_dispersion, x = model)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Seed dispersion (m)") +
  theme(axis.title.x = element_blank()) -> b

boxplots <- plot_grid(a,b, ncol = 2, align = "h", labels = c("b", "c"))

plot_grid(histograms, boxplots, nrow=2, labels = c("a", ""))

# Edit redundant labels
# Fix names for axes

## Long distance dispersal

The mean dispersal in the boxplots is not the same as the dispersal distances before. This mean dispersal is the distance from the origin to the mean seed location in each simulation run. The mean seed location is used to calculate dispersion, as the average distance of each seed to the mean seed location, as shown in Figure 1.

```{r}
knitr::kable(dispersal_kernel_table %>% 
               rename(Long_distance = LDD))
```

## Weibull seed dispersal kernel 
```{r}
knitr::kable(weibull_table)
```
